---
title: "Rhetorics Analyses From Cleaned Data"
author: "Hannah Hok Kim"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  warning: false
  message: false
params:
  data_d: "../Data"
---

# Setup

```{r}
required_packages <- c(
  "tidyverse",
  "psych",
  "broom",
  "knitr",
  "scales",
  "patchwork"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0) {
  stop(
    "Install missing package(s) before running this notebook: ",
    paste(missing_packages, collapse = ", ")
  )
}

library(tidyverse)
library(psych)
library(broom)
library(knitr)
library(patchwork)

theme_set(theme_bw())

clean <- readRDS(file.path(params$data_d, "responses_final_complete_clean.rds"))

practice_labels <- c(
  datashare = "Data sharing",
  prereg = "Preregistration",
  OA = "Open access",
  replication = "Replication"
)

field_order <- c(
  "Economics",
  "Business",
  "Political Science",
  "Psychology",
  "Communications",
  "Sociology",
  "Education",
  "Anthropology"
)

as_score <- function(x) {
  if (is.factor(x)) {
    as.numeric(x)
  } else {
    as.numeric(x)
  }
}

as_binary_yes <- function(x) {
  out <- case_when(
    as.character(x) == "Yes" ~ 1,
    as.character(x) == "No" ~ 0,
    TRUE ~ NA_real_
  )
  out
}

round_numeric <- function(data, digits = 2) {
  data %>%
    mutate(across(where(is.numeric), ~ round(.x, digits)))
}

item_score_df <- clean %>%
  transmute(
    across(c(pos1:pos5, inc1, inc2, inc3_r, inc4_r, neut1_r, neut2, neut3_r, neut4, neut5), as_score)
  )

analysis_df <- clean %>%
  mutate(
    field = factor(field_first, levels = field_order),
    n_fields = lengths(field_multi),
    field_multi_group = if_else(n_fields == 1, as.character(field), "Multiple"),
    qualquant = fct_drop(qualquant),
    years_clean = if_else(years > 80, NA_real_, years),
    across(
      c(
        datashare_opinion,
        prereg_opinion,
        OA_opinion,
        replication_opinion
      ),
      as_score,
      .names = "{.col}_score"
    ),
    across(
      c(
        datashare_behavior,
        prereg_behavior,
        OA_behavior,
        replication_behavior
      ),
      as_binary_yes,
      .names = "{.col}_binary"
    )
  )

missing_field_n <- sum(is.na(analysis_df$field))
field_analysis_df <- analysis_df %>%
  filter(!is.na(field))

attitude_outcomes <- c(
  "datashare_opinion_score",
  "prereg_opinion_score",
  "OA_opinion_score",
  "replication_opinion_score"
)

behavior_outcomes <- c(
  "datashare_behavior_binary",
  "prereg_behavior_binary",
  "OA_behavior_binary",
  "replication_behavior_binary"
)

granular_outcomes <- c(
  "datashare_b_percent_1",
  "prereg_b_percent_1",
  "OA_b_percent_1",
  "replication_b_percent_1",
  "datashare_b_others_1",
  "prereg_b_others_1",
  "OA_b_others_1",
  "replication_b_others_1",
  "OA_b_when_1",
  "replication_b_when_1"
)

outcome_labels <- c(
  datashare_opinion_score = "Data sharing attitude",
  prereg_opinion_score = "Preregistration attitude",
  OA_opinion_score = "Open access attitude",
  replication_opinion_score = "Replication attitude",
  datashare_behavior_binary = "Data sharing behavior",
  prereg_behavior_binary = "Preregistration behavior",
  OA_behavior_binary = "Open access behavior",
  replication_behavior_binary = "Replication behavior",
  datashare_b_percent_1 = "Data sharing: percent of work",
  prereg_b_percent_1 = "Preregistration: percent of work",
  OA_b_percent_1 = "Open access: percent of work",
  replication_b_percent_1 = "Replication: percent of work",
  datashare_b_others_1 = "Data sharing: perceived others",
  prereg_b_others_1 = "Preregistration: perceived others",
  OA_b_others_1 = "Open access: perceived others",
  replication_b_others_1 = "Replication: perceived others",
  OA_b_when_1 = "Open access: timing",
  replication_b_when_1 = "Replication: timing"
)
```

# Sample Characteristics

```{r}
sample_n <- nrow(analysis_df)

field_counts <- field_analysis_df %>%
  count(field, sort = TRUE)

method_counts <- analysis_df %>%
  count(qualquant, sort = TRUE, name = "n") %>%
  mutate(qualquant = fct_na_value_to_level(qualquant, level = "Missing"))

years_summary <- analysis_df %>%
  summarise(
    n = sum(!is.na(years_clean)),
    mean = mean(years_clean, na.rm = TRUE),
    sd = sd(years_clean, na.rm = TRUE),
    median = median(years_clean, na.rm = TRUE),
    min = min(years_clean, na.rm = TRUE),
    max = max(years_clean, na.rm = TRUE)
  )

sample_n
kable(field_counts, caption = "Analytic sample size by field")
kable(method_counts, caption = "Analytic sample size by methodological orientation")
kable(round_numeric(years_summary, 2), caption = "Years of research experience")
```

```{r}
ggplot(field_counts, aes(x = fct_reorder(field, n), y = n)) +
  geom_col(fill = "#4E79A7") +
  coord_flip() +
  labs(x = NULL, y = "Respondents", title = "Sample size by field")
```

```{r}
ggplot(analysis_df, aes(x = years_clean)) +
  geom_histogram(binwidth = 2, boundary = 0, fill = "#59A14F", color = "white") +
  labs(x = "Years of research experience", y = "Respondents")
```

# Scale Reliability

Neutrality is included in descriptive analyses, but confirmatory models use Positivism and Inclusivity only.

```{r}
scale_reliability <- tibble(
  scale = c("Positivism", "Inclusivity", "Neutrality"),
  alpha = c(
    psych::alpha(item_score_df[paste0("pos", 1:5)])$total$raw_alpha,
    psych::alpha(item_score_df[c("inc1", "inc2", "inc3_r", "inc4_r")])$total$raw_alpha,
    psych::alpha(item_score_df[c("neut1_r", "neut2", "neut3_r", "neut4", "neut5")])$total$raw_alpha
  )
) %>%
  mutate(alpha = round(alpha, 3))

kable(scale_reliability, caption = "Internal consistency of epistemological belief scales")
```

# H1: Epistemological Beliefs By Field

```{r}
epistemology_by_field <- field_analysis_df %>%
  group_by(field) %>%
  summarise(
    n = n(),
    Positivism_M = mean(Positivism, na.rm = TRUE),
    Positivism_SD = sd(Positivism, na.rm = TRUE),
    Inclusivity_M = mean(Inclusivity, na.rm = TRUE),
    Inclusivity_SD = sd(Inclusivity, na.rm = TRUE),
    Neutrality_M = mean(Neutrality, na.rm = TRUE),
    Neutrality_SD = sd(Neutrality, na.rm = TRUE),
    .groups = "drop"
  )

kable(round_numeric(epistemology_by_field, 2), caption = "Epistemological beliefs by field")
```

```{r}
epistemology_long <- field_analysis_df %>%
  select(ResponseId, field, Positivism, Inclusivity, Neutrality) %>%
  pivot_longer(
    cols = c(Positivism, Inclusivity, Neutrality),
    names_to = "scale",
    values_to = "score"
  )

ggplot(epistemology_long, aes(x = field, y = score, fill = field)) +
  geom_boxplot(outlier.alpha = 0.15) +
  stat_summary(fun = mean, geom = "point", shape = 23, fill = "white", size = 2.5) +
  coord_flip() +
  facet_wrap(~ scale, nrow = 1) +
  guides(fill = "none") +
  labs(x = NULL, y = "Scale score")
```

```{r}
field_model_tests <- epistemology_long %>%
  group_by(scale) %>%
  group_modify(~ {
    fit <- lm(score ~ field, data = .x)
    broom::tidy(anova(fit)) %>%
      filter(term == "field")
  }) %>%
  ungroup() %>%
  mutate(p_bh = p.adjust(p.value, method = "BH"))

kable(field_model_tests, caption = "Field differences in epistemological beliefs")
```

# H2: Epistemological Beliefs By Methodological Orientation

```{r}
method_df <- analysis_df %>%
  filter(!is.na(qualquant))

epistemology_by_method <- method_df %>%
  group_by(qualquant) %>%
  summarise(
    n = n(),
    Positivism_M = mean(Positivism, na.rm = TRUE),
    Positivism_SD = sd(Positivism, na.rm = TRUE),
    Inclusivity_M = mean(Inclusivity, na.rm = TRUE),
    Inclusivity_SD = sd(Inclusivity, na.rm = TRUE),
    Neutrality_M = mean(Neutrality, na.rm = TRUE),
    Neutrality_SD = sd(Neutrality, na.rm = TRUE),
    .groups = "drop"
  )

kable(round_numeric(epistemology_by_method, 2), caption = "Epistemological beliefs by methodological orientation")
```

```{r}
method_long <- method_df %>%
  select(ResponseId, qualquant, Positivism, Inclusivity, Neutrality) %>%
  pivot_longer(
    cols = c(Positivism, Inclusivity, Neutrality),
    names_to = "scale",
    values_to = "score"
  )

ggplot(method_long, aes(x = qualquant, y = score, fill = qualquant)) +
  geom_boxplot(outlier.alpha = 0.15) +
  stat_summary(fun = mean, geom = "point", shape = 23, fill = "white", size = 2.5) +
  facet_wrap(~ scale, nrow = 1) +
  guides(fill = "none") +
  labs(x = "Methodological orientation", y = "Scale score")
```

```{r}
method_model_tests <- method_long %>%
  group_by(scale) %>%
  group_modify(~ {
    fit <- lm(score ~ qualquant, data = .x)
    broom::tidy(anova(fit)) %>%
      filter(term == "qualquant")
  }) %>%
  ungroup() %>%
  mutate(p_bh = p.adjust(p.value, method = "BH"))

kable(method_model_tests, caption = "Methodological-orientation differences in epistemological beliefs")
```

# Perceived Methodological Composition Of Field

```{r}
correlation_table <- function(data, predictor, outcomes) {
  map_dfr(outcomes, function(outcome) {
    dat <- data %>%
      select(all_of(c(predictor, outcome))) %>%
      drop_na()

    test <- cor.test(dat[[predictor]], dat[[outcome]])

    tibble(
      outcome = outcome,
      n = nrow(dat),
      r = unname(test$estimate),
      conf_low = test$conf.int[1],
      conf_high = test$conf.int[2],
      p = test$p.value
    )
  }) %>%
    mutate(
      outcome = recode(outcome, !!!outcome_labels),
      p_bh = p.adjust(p, method = "BH")
    )
}

belief_composition_correlations <- correlation_table(
  analysis_df,
  predictor = "qualquantField_1",
  outcomes = c("Positivism", "Inclusivity", "Neutrality")
)

kable(
  round_numeric(belief_composition_correlations, 3),
  caption = "Associations between perceived qualitative composition of field and epistemological beliefs"
)
```

```{r}
analysis_df %>%
  select(qualquantField_1, Positivism, Inclusivity, Neutrality) %>%
  pivot_longer(
    cols = c(Positivism, Inclusivity, Neutrality),
    names_to = "scale",
    values_to = "score"
  ) %>%
  ggplot(aes(x = qualquantField_1, y = score)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_wrap(~ scale, nrow = 1) +
  labs(
    x = "Estimated percent qualitative researchers in field",
    y = "Epistemological belief score"
  )
```

# H3: Open Science Attitudes And Behaviors

```{r}
attitude_long <- field_analysis_df %>%
  select(ResponseId, field, all_of(attitude_outcomes)) %>%
  pivot_longer(
    cols = all_of(attitude_outcomes),
    names_to = "outcome",
    values_to = "score"
  ) %>%
  mutate(
    practice = recode(outcome, !!!outcome_labels)
  )

behavior_long <- field_analysis_df %>%
  select(ResponseId, field, all_of(behavior_outcomes)) %>%
  pivot_longer(
    cols = all_of(behavior_outcomes),
    names_to = "outcome",
    values_to = "behavior"
  ) %>%
  mutate(
    practice = recode(outcome, !!!outcome_labels)
  )

attitude_summary <- attitude_long %>%
  group_by(field, practice) %>%
  summarise(
    n = sum(!is.na(score)),
    mean = mean(score, na.rm = TRUE),
    sd = sd(score, na.rm = TRUE),
    se = sd / sqrt(n),
    .groups = "drop"
  )

behavior_summary <- behavior_long %>%
  group_by(field, practice) %>%
  summarise(
    n = sum(!is.na(behavior)),
    prop = mean(behavior, na.rm = TRUE),
    se = sqrt(prop * (1 - prop) / n),
    .groups = "drop"
  )

kable(round_numeric(attitude_summary, 2), caption = "Open science attitudes by field")
kable(round_numeric(behavior_summary, 3), caption = "Open science behaviors by field")
```

```{r}
ggplot(attitude_summary, aes(x = mean, y = field, color = practice)) +
  geom_point(position = position_dodge(width = 0.5), size = 2.5) +
  geom_errorbar(
    aes(xmin = mean - se, xmax = mean + se),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  scale_x_continuous(limits = c(1, 5)) +
  labs(x = "Mean attitude score", y = NULL, color = "Outcome")
```

```{r}
ggplot(behavior_summary, aes(x = prop, y = field, color = practice)) +
  geom_point(position = position_dodge(width = 0.5), size = 2.5) +
  geom_errorbar(
    aes(xmin = pmax(prop - se, 0), xmax = pmin(prop + se, 1)),
    position = position_dodge(width = 0.5),
    width = 0.2
  ) +
  scale_x_continuous(labels = scales::percent_format()) +
  labs(x = "Percent reporting behavior", y = NULL, color = "Outcome")
```

# More Granular Behavioral Measures

These analyses address Micah's note that binary ever/never measures may miss variation in how strongly, regularly, or centrally researchers engage in open science practices.

```{r}
granular_existing <- granular_outcomes[granular_outcomes %in% names(analysis_df)]

granular_summary <- analysis_df %>%
  select(ResponseId, field, all_of(granular_existing)) %>%
  pivot_longer(
    cols = all_of(granular_existing),
    names_to = "outcome",
    values_to = "value"
  ) %>%
  mutate(outcome = recode(outcome, !!!outcome_labels)) %>%
  group_by(outcome) %>%
  summarise(
    n = sum(!is.na(value)),
    mean = mean(value, na.rm = TRUE),
    sd = sd(value, na.rm = TRUE),
    median = median(value, na.rm = TRUE),
    min = min(value, na.rm = TRUE),
    max = max(value, na.rm = TRUE),
    .groups = "drop"
  )

kable(round_numeric(granular_summary, 2), caption = "Granular open science behavior measures")
```

```{r}
granular_correlations <- map_dfr(granular_existing, function(outcome) {
  correlation_table(
    analysis_df,
    predictor = "Positivism",
    outcomes = outcome
  ) %>%
    mutate(predictor = "Positivism") %>%
    bind_rows(
      correlation_table(
        analysis_df,
        predictor = "Inclusivity",
        outcomes = outcome
      ) %>%
        mutate(predictor = "Inclusivity")
    )
}) %>%
  select(outcome, predictor, n, r, conf_low, conf_high, p, p_bh)

kable(
  round_numeric(granular_correlations, 3),
  caption = "Associations between epistemological beliefs and granular behavioral measures"
)
```

# Confirmatory Nested Models

Confirmatory models exclude Neutrality because its reliability is below the desired threshold.

Model 1 includes field. Model 2 adds methodological orientation. Model 3 adds Positivism and Inclusivity.

```{r}
make_model_data <- function(data, outcome) {
  data %>%
    select(
      all_of(outcome),
      field,
      qualquant,
      Positivism,
      Inclusivity
    ) %>%
    drop_na()
}

fit_attitude_models <- function(data, outcome) {
  dat <- make_model_data(data, outcome)

  m1 <- lm(reformulate("field", response = outcome), data = dat)
  m2 <- lm(reformulate(c("field", "qualquant"), response = outcome), data = dat)
  m3 <- lm(
    reformulate(c("field", "qualquant", "Positivism", "Inclusivity"), response = outcome),
    data = dat
  )

  list(
    outcome = outcome,
    n = nrow(dat),
    model1 = m1,
    model2 = m2,
    model3 = m3,
    compare_12 = anova(m1, m2),
    compare_23 = anova(m2, m3),
    compare_13 = anova(m1, m3)
  )
}

fit_behavior_models <- function(data, outcome) {
  dat <- make_model_data(data, outcome)

  m1 <- glm(reformulate("field", response = outcome), data = dat, family = binomial)
  m2 <- glm(reformulate(c("field", "qualquant"), response = outcome), data = dat, family = binomial)
  m3 <- glm(
    reformulate(c("field", "qualquant", "Positivism", "Inclusivity"), response = outcome),
    data = dat,
    family = binomial
  )

  list(
    outcome = outcome,
    n = nrow(dat),
    model1 = m1,
    model2 = m2,
    model3 = m3,
    compare_12 = anova(m1, m2, test = "Chisq"),
    compare_23 = anova(m2, m3, test = "Chisq"),
    compare_13 = anova(m1, m3, test = "Chisq")
  )
}

attitude_models <- set_names(
  map(attitude_outcomes, ~ fit_attitude_models(analysis_df, .x)),
  attitude_outcomes
)

behavior_models <- set_names(
  map(behavior_outcomes, ~ fit_behavior_models(analysis_df, .x)),
  behavior_outcomes
)
```

```{r}
attitude_model_comparisons <- map_dfr(attitude_models, function(res) {
  tibble(
    outcome = res$outcome,
    n = res$n,
    model1_adj_r2 = summary(res$model1)$adj.r.squared,
    model2_adj_r2 = summary(res$model2)$adj.r.squared,
    model3_adj_r2 = summary(res$model3)$adj.r.squared,
    model1_aic = AIC(res$model1),
    model2_aic = AIC(res$model2),
    model3_aic = AIC(res$model3),
    p_12 = res$compare_12$`Pr(>F)`[2],
    p_23 = res$compare_23$`Pr(>F)`[2],
    p_13 = res$compare_13$`Pr(>F)`[2]
  )
}) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_labels),
    p_12_bh = p.adjust(p_12, method = "BH"),
    p_23_bh = p.adjust(p_23, method = "BH"),
    p_13_bh = p.adjust(p_13, method = "BH")
  )

behavior_model_comparisons <- map_dfr(behavior_models, function(res) {
  tibble(
    outcome = res$outcome,
    n = res$n,
    model1_aic = AIC(res$model1),
    model2_aic = AIC(res$model2),
    model3_aic = AIC(res$model3),
    p_12 = res$compare_12$`Pr(>Chi)`[2],
    p_23 = res$compare_23$`Pr(>Chi)`[2],
    p_13 = res$compare_13$`Pr(>Chi)`[2]
  )
}) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_labels),
    p_12_bh = p.adjust(p_12, method = "BH"),
    p_23_bh = p.adjust(p_23, method = "BH"),
    p_13_bh = p.adjust(p_13, method = "BH")
  )

kable(round_numeric(attitude_model_comparisons, 3), caption = "Nested model comparisons for attitudinal outcomes")
kable(round_numeric(behavior_model_comparisons, 3), caption = "Nested model comparisons for behavioral outcomes")
```

```{r}
attitude_epistemology_coefficients <- map_dfr(attitude_models, function(res) {
  broom::tidy(res$model3, conf.int = TRUE) %>%
    filter(term %in% c("Positivism", "Inclusivity")) %>%
    mutate(
      outcome = res$outcome,
      n = res$n
    )
}) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_labels),
    p_bh = p.adjust(p.value, method = "BH")
  ) %>%
  select(outcome, n, term, estimate, conf.low, conf.high, p.value, p_bh)

behavior_epistemology_coefficients <- map_dfr(behavior_models, function(res) {
  broom::tidy(res$model3, conf.int = TRUE) %>%
    filter(term %in% c("Positivism", "Inclusivity")) %>%
    mutate(
      outcome = res$outcome,
      n = res$n,
      odds_ratio = exp(estimate),
      odds_ratio_low = exp(conf.low),
      odds_ratio_high = exp(conf.high)
    )
}) %>%
  mutate(
    outcome = recode(outcome, !!!outcome_labels),
    p_bh = p.adjust(p.value, method = "BH")
  ) %>%
  select(
    outcome,
    n,
    term,
    estimate,
    conf.low,
    conf.high,
    odds_ratio,
    odds_ratio_low,
    odds_ratio_high,
    p.value,
    p_bh
  )

kable(round_numeric(attitude_epistemology_coefficients, 3), caption = "Epistemological belief coefficients from full attitude models")
kable(round_numeric(behavior_epistemology_coefficients, 3), caption = "Epistemological belief coefficients from full behavior models")
```

```{r}
attitude_fit_plot <- attitude_model_comparisons %>%
  select(outcome, model1_adj_r2, model2_adj_r2, model3_adj_r2) %>%
  pivot_longer(
    cols = starts_with("model"),
    names_to = "model",
    values_to = "adj_r2"
  ) %>%
  mutate(
    model = recode(
      model,
      model1_adj_r2 = "Model 1: field",
      model2_adj_r2 = "Model 2: + method",
      model3_adj_r2 = "Model 3: + beliefs"
    ),
    model = factor(model, levels = c("Model 1: field", "Model 2: + method", "Model 3: + beliefs"))
  )

behavior_fit_plot <- behavior_model_comparisons %>%
  select(outcome, model1_aic, model2_aic, model3_aic) %>%
  pivot_longer(
    cols = starts_with("model"),
    names_to = "model",
    values_to = "aic"
  ) %>%
  mutate(
    model = recode(
      model,
      model1_aic = "Model 1: field",
      model2_aic = "Model 2: + method",
      model3_aic = "Model 3: + beliefs"
    ),
    model = factor(model, levels = c("Model 1: field", "Model 2: + method", "Model 3: + beliefs"))
  )

p_attitude_fit <- ggplot(attitude_fit_plot, aes(x = model, y = adj_r2, group = outcome, color = outcome)) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = NULL, y = "Adjusted R-squared", color = "Outcome")

p_behavior_fit <- ggplot(behavior_fit_plot, aes(x = model, y = aic, group = outcome, color = outcome)) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = NULL, y = "AIC", color = "Outcome")

p_attitude_fit / p_behavior_fit
```

# Manuscript Number Check

Use this section to keep the prose synchronized with the current cleaned file.

```{r}
current_numbers <- list(
  analytic_n = nrow(analysis_df),
  field_counts = field_counts,
  method_counts = method_counts,
  years_summary = years_summary,
  reliability = scale_reliability,
  attitude_model_comparisons = attitude_model_comparisons,
  behavior_model_comparisons = behavior_model_comparisons
)

current_numbers
```
